Gravitational Lens Modelling of Flux Ratio Anomalous System B1555+375

### 2015-04-22 papers/B1555_EdgeOn/B1555_letter.tex
More small edits to the text

While future large-scale surveys, such as the LSST, are expected to
deliver new large samples of gravitationally lensed quasars, in the
near future the discovery of new systems using narrow-line quasar
emission \citep{N14} will provide new invaluable statistical
perspective on current constraints based on flux ratio anomalies.  In
this paper, we explore the idea that not all flux ratio anomalies are
related to the presence of substructure, but that some originate in a
more complex lens galaxy mass distribution than initially
considered. To this end we use new infra-red data from the Strong
lensing at High Angular Resolution Program (SHARP; Fassnacht et al.,
in prep). The main goal of the SHARP project is to re-observe known
quadruple and Einstein ring lens systems at a much higher angular
resolution by making use of observations with (1) adaptive optics (AO)
on the Keck telescopes, (2) the Hubble Space Telescope, and (3) large
interferometric arrays \citep{SHARP12,V12}. In particular, to better
understand the causes of flux ratio anomalies in gravitational lenses
we focus on the B1555+375 lens system\citep{Marlow}.  This system was
discovered in the Cosmic Lens All-Sky Survey \citep{CLASS1,CLASS2} and
has a pair of merging images that show a strong flux ratio anomaly at
radio wavelengths.  Follow-up observations ruled out radio propagation
effects as the cause of the anomaly \citep{K03,KD04}.  The B1555+375
system was included in both the observational and theoretical analysis
of \citet{Dalal2002} and \citet{Xu14} and, thus, understanding its
mass distribution is of key importantce.  Our analysis makes use of
the new information obtained from the latest SHARP infrared imaging.
The paper is organized as follows: the latest SHARP data are
introduced in section 2; in section 3 we present the lens model while
the results of this analysis and their implications are discussed in
section 4.

@@

The B1555+375 system was observed using the NIRC2 camera on the Keck 2
Telescope on the night of 2012 May 16 UT.  The adaptive optics system
was used, with the corrections derived from the laser guide star and a
$R$=14.4 tip-tilt that was located 45$^{\prime\prime}$ from the lens
system.  The ``narrow camera'' mode was used, giving a field of view
of roughly 10$^{\prime\prime}$\ on a side and a pixel scale of 10~mas.
Six dithered 300~s exposures were obtained in $K^{\prime}$ band.  The
data were reduced with the standard SHARP pipeline, which is a
python-based package that is a refinement of the process described in
\citet{Auger_EELS1}.  A cutout of the final reduced image is shown in
Fig.~\ref{fig:multiband} and again in Fig.~\ref{fig:merlin} with
contours from the MERLIN radio observations of \citet{Marlow}
overlaid.  There are several notable features in the Keck AO data.
These include the nearly complete lack of emission associated with the
B and D lensed images, as can be seen by comparing the radio contours
to the $K^\prime$-band emission, and faint but clearly visible
emission from what appears to be an edge-on lensing galaxy.

@@

To model B1555+375, we use the lens modeling code {\tt glafic}
\citep{Oguri}.  The inputs to the model are the observed image
positions and flux densities measured by the radio observations of
\citet{Marlow}.  These authors present a singular isothermal ellipsoid
(SIE) model that provides the first fit to the image positions but
does not reproduce the strong flux ratio anomaly between components A
and B.  \citep[see Fig.~6 and Tables 2 \& 3 in][]{Marlow}.  Although
the flux ratio anomaly that is observed in the radio data may be
caused by substructure in the lensing galaxy, our high-resolution near
infrared imaging suggests another explanation.  In both the $K^\prime$
band adaptive optics data and the HST imaging (two rightmost panels in
Fig. 1), the lensing galaxy has a clear edge-on disk component with a
position angle of $\approx 10$ degree.  We therefore model the system
with a disk component in order to see if such a model can explain the
flux ratio anomaly without the need for substructure.  The edge-on
disk may also explain the system morphology that is observed in the
near infrared imaging.  In particular, neither image B nor image D is
detected in the Keck AO data (Fig.~\ref{fig:merlin}), and both of
these lensed images lie on or close to the edge-on disk.  Thus, strong
extinction in the disk could cause the lack of detection of these
images.

As a first trail of modelling, we create a single SIE model to 
justify the perforamnce of a simple lensing potential. Consistent with previous studies
\citep{Marlow, Xu14}, our single elliptical potential model cannot predict flux anomaly in B1555+375. 
We then create two separate models of the lensing potential.  In the
first we treat the lens as a SIE bulge and an exponential edge-on disk
(SIE+expdisk).  In the second model the disk is represented as a very
elongated SIE component (2-SIE).  The lens system is red, with a
strong jump in flux between the F814W (roughly I) and F160W (roughly
H) bands (Fig.~\ref{fig:multiband}), so we assign a lens redshift of
$z_\ell = 1.0$ and a source redshift of $z_s =1.5$. Our steps to explore the parameter space are: 1) Apply strong priors from SHARP AO image to the edge-on disk position angle and ellipticity, and use the best-fit model of \citet{marlow} as the start point of the bulge component and the source position. 2) Let all parameters without priors vary to reach a temporary best-fit. 3) Start with the best-fit in step 2 and let all parameters with priors vary to reach another best-fit (also ensure this best-fit is not far from the priors). 4) Let all parameters vary to reach the final optimal model. 
The full best-fit parameters are shown in Table
2. Both of these models predict fairly matched results with the radio
observation, and the 2-SIE model has a better performance in flux
ratios. The comparison of observation and our lens modelling is given
in Figure 3 and Tables 3 \& 4.
[WHY NO EXTERNAL SHEAR?---WE THINK WITH ONLY RADIO DATA ITS LACK OF CONSTRAINT]

@@

\begin{figure}
%\plotone{1555_ao_merlin_overlay.eps}
\includegraphics[width=84mm]{1555_ao_merlin_overlay.eps}
\caption{High-resolution  $K^\prime$ band imaging of the B1555+375 lens system 
with contours from the MERLIN observation of \citet{Marlow} overlaid.
%
\label{fig:merlin}}
\end{figure}

@@

Comparision between radio and model-predicted lensed image positions. Observation data quoted from Table 2 in \citet{Marlow}. Position offsets are in unit of arcsec. Both SIE+expdisk and 2-SIE lens model-predicted positions well match to the radio observation.

@@

%\end{minipage}
\medskip
Comparision between radio data and model-predicted flux ratios. The radio observation data is quoted from Tables 2 and Table 3 in \citet{Marlow}. 1-SIE model is quoted from Tabel 5 in \citet{Marlow}.

@@

In our models based on the new imaging from the SHARP $K^{\prime}$
band observations, the most significant difference in comparison to
previous modelling is a new edge-on disk component. With a bulge plus
an edge-on disk, our models produce matches to both the radio image
positions and the flux ratio, while models with only one SIE
component \citep{Marlow,Xu14} fail to reproduce the fluxes
correctly. Our successful modelling indicates that the
kiloparsec-scale structure of lens galaxy itself is sufficient to
explain the flux ratio anomalies in B1555+375, without any need to
invoke the presense of substructure.

>>>>

While future large-scale surveys, such as the LSST, are expected to
deliver new large samples of gravitationally lensed quasars, in the
near future the discovery of new systems using narrow-line quasar
emission \citep{N14} will provide new invaluable statistical
perspective on current constraints based on flux ratio anomalies.  In
this paper, we explore the idea that not all flux ratio anomalies are
related to the presence of substructure, but that some originate in a
more complex lens galaxy mass distribution than initially
considered. To this end we use new infra-red data from the Strong
lensing at High Angular Resolution Program (SHARP; Fassnacht et al.,
in prep). The main goal of the SHARP project is to re-observe known
quadruple and Einstein ring lens systems at a much higher angular
resolution by making use of observations with (1) adaptive optics (AO)
on the Keck telescopes, (2) the Hubble Space Telescope, and (3) large
interferometric arrays \citep{SHARP12,V12}. In particular, to better
understand the causes of flux ratio anomalies in gravitational lenses
we focus on the B1555+375 lens system\citep{Marlow99}.  This system was
discovered in the Cosmic Lens All-Sky Survey \citep{CLASS1,CLASS2} and
has a pair of merging images that show a strong flux ratio anomaly at
radio wavelengths.  Follow-up observations ruled out radio propagation
effects as the cause of the anomaly \citep{K03,KD04}.  The B1555+375
system was included in both the observational and theoretical analysis
of \citet{Dalal2002} and \citet{Xu14} and, thus, understanding its
mass distribution is of key importantce.  Our analysis makes use of
the new information obtained from the latest SHARP infrared imaging.
The paper is organized as follows: the latest SHARP data are
introduced in section 2; in section 3 we present the lens model while
the results of this analysis and their implications are discussed in
section 4.

@@

The B1555+375 system was observed using the NIRC2 camera on the Keck 2
Telescope on the night of 2012 May 16 UT.  The adaptive optics system
was used, with the corrections derived from the laser guide star and a
$R$=14.4 tip-tilt that was located 45$^{\prime\prime}$ from the lens
system.  The ``narrow camera'' mode was used, giving a field of view
of roughly 10$^{\prime\prime}$\ on a side and a pixel scale of 10~mas.
Six dithered 300~s exposures were obtained in $K^{\prime}$ band.  The
data were reduced with the standard SHARP pipeline, which is a
python-based package that is a refinement of the process described in
\citet{Auger_EELS1}.  A cutout of the final reduced image is shown in
Fig.~\ref{fig:multiband} and again in Fig.~\ref{fig:merlin} with
contours from the MERLIN radio observations of \citet{Marlow99}
overlaid.  There are several notable features in the Keck AO data.
These include the nearly complete lack of emission associated with the
B and D lensed images, as can be seen by comparing the radio contours
to the $K^\prime$-band emission, and faint but clearly visible
emission from what appears to be an edge-on lensing galaxy.

@@

To model B1555+375, we use the lens modeling code {\tt glafic}
\citep{Oguri}.  The inputs to the model are the observed image
positions and flux densities measured by the radio observations of
\citet{Marlow99}.  These authors present a singular isothermal ellipsoid
(SIE) model that provides the first fit to the image positions but
does not reproduce the strong flux ratio anomaly between components A
and B.  \citep[see Fig.~6 and Tables 2 \& 3 in][]{Marlow99}.  Although
the flux ratio anomaly that is observed in the radio data may be
caused by substructure in the lensing galaxy, our high-resolution near
infrared imaging suggests another explanation.  In both the $K^\prime$
band adaptive optics data and the HST imaging (two rightmost panels in
Fig. 1), the lensing galaxy has a clear edge-on disk component with a
position angle of $\approx 10$ degree.  We therefore model the system
with a disk component in order to see if such a model can explain the
flux ratio anomaly without the need for substructure.  The edge-on
disk may also explain the system morphology that is observed in the
near infrared imaging.  In particular, neither image B nor image D is
detected in the Keck AO data (Fig.~\ref{fig:merlin}), and both of
these lensed images lie on or close to the edge-on disk.  Thus, strong
extinction in the disk could cause the lack of detection of these
images.

As a first trail of modelling, we create a single SIE model to 
justify the perforamnce of a simple lensing potential. Consistent with previous studies
\citep{Marlow99, Xu14}, our single elliptical potential model cannot predict flux anomaly in B1555+375. 
We then create two separate models of the lensing potential.  In the
first we treat the lens as a SIE bulge and an exponential edge-on disk
(SIE+expdisk).  In the second model the disk is represented as a very
elongated SIE component (2-SIE).  The lens system is red, with a
strong jump in flux between the F814W (roughly I) and F160W (roughly
H) bands (Fig.~\ref{fig:multiband}), so we assign a lens redshift of
$z_\ell = 1.0$ and a source redshift of $z_s =1.5$. Our steps to explore the parameter space are: 1) Apply strong priors from SHARP AO image to the edge-on disk position angle and ellipticity, and use the best-fit model of \citet{Marlow99} as the start point of the bulge component and the source position. 2) Let all parameters without priors vary to reach a temporary best-fit. 3) Start with the best-fit in step 2 and let all parameters with priors vary to reach another best-fit (also ensure this best-fit is not far from the priors). 4) Let all parameters vary to reach the final optimal model. 
The full best-fit parameters are shown in Table
2. Both of these models predict fairly matched results with the radio
observation, and the 2-SIE model has a better performance in flux
ratios. The comparison of observation and our lens modelling is given
in Figure 3 and Tables 3 \& 4.
[WHY NO EXTERNAL SHEAR?---WITH ONLY THE RADIO POSITIONS AND FLUXES AS
CONSTRAINTS, ADDING SHEAR WOULD MAKE THE PROBLEM EVEN MORE UNDERCONSTRAINED]

@@

\begin{figure}
%\plotone{1555_ao_merlin_overlay.eps}
\includegraphics[width=84mm]{1555_ao_merlin_overlay.eps}
\caption{High-resolution  $K^\prime$ band imaging of the B1555+375 lens system 
with contours from the MERLIN observation of \citet{Marlow99} overlaid.
%
\label{fig:merlin}}
\end{figure}

@@

Comparision between radio and model-predicted lensed image positions. Observation data quoted from Table 2 in \citet{Marlow99}. Position offsets are in unit of arcsec. Both SIE+expdisk and 2-SIE lens model-predicted positions well match to the radio observation.

@@

%\end{minipage}
\medskip
Comparision between radio data and model-predicted flux ratios. The radio observation data is quoted from Tables 2 and Table 3 in \citet{Marlow99}. 1-SIE model is quoted from Tabel 5 in \citet{Marlow99}.

@@

In our models based on the new imaging from the SHARP $K^{\prime}$
band observations, the most significant difference in comparison to
previous modelling is a new edge-on disk component. With a bulge plus
an edge-on disk, our models produce matches to both the radio image
positions and the flux ratio, while models with only one SIE
component \citep{Marlow99,Xu14} fail to reproduce the fluxes
correctly. Our successful modelling indicates that the
kiloparsec-scale structure of lens galaxy itself is sufficient to
explain the flux ratio anomalies in B1555+375, without any need to
invoke the presense of substructure.
